Spectral partitioning identifies individual heterogeneity in the functional network topography of ventral and anterior medial prefrontal

### 2019-09-13 Manuscript.Rmd
intro rehashed

---
title: Individual heterogeneity in the functional network topography of medial prefrontal
  cortex
author: "Claudio Toro-Serey^1^, Sean M. Tobyne^1,2^, & Joseph T. McGuire^1^"
csl: apa.csl
output:
  word_document: #default
    reference_docx: Style_reference_draft.docx
  pdf_document:
    highlight: haddock
    keep_tex: no
header-includes:
- \usepackage{setspace}\doublespacing
- \usepackage[left]{lineno}
- \linenumbers
- \usepackage{float}
- \newcommand{\beginsupplement}{ \setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}}
  \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}} }
#subtitle: 'Abbreviated Title: Heterogeneity in the functional topography of mPFC'
bibliography: mPFC_paper.bib
---

@@

1.  Department of Psychological and Brain Sciences, Boston University, Boston, USA

@@

  Human medial prefrontal cortex (mPFC) and posterior cingulate cortex (PCC) are implicated in multiple cognitive functions, including subjective valuation processes and processes linked to the default network (DN). Our ability to interpret these seemingly co-localized effects is constrained by a limited understanding of the individual-level heterogeneity in mPFC/PCC functional organization. Here we used cortical surface-based meta-analysis to identify a parcel in human PCC that was preferentially implicated in DN effects relative to valuation. We then used resting-state fMRI data and a data-driven network analysis algorithm, spectral partitioning, to partition mPFC and PCC into "DN" and "non-DN" subdivisions in individual participants (n = 100 from the Human Connectome Project). The spectral partitioning algorithm identified individual-level cortical subdivisions that were reliable across test/retest datasets and varied markedly across individuals, especially in mPFC. Location of the DN was informed by cortical folding, and resembled a recently proposed fractionation of this network. Our results point toward a new generation of strategies for assessing whether distinct cognitive functions engage common or distinct mPFC subregions, especially at the individual level.

@@

  Human medial prefrontal cortex (mPFC) and posterior cingulate cortex (PCC) are jointly associated with a large and diverse set of cognitive processes [@Hiser2018; @Kragel2018], and are among the regions implicated in a set of functions associated with the default network (DN) [@Buckner2008]. In humans, areas in the DN are chiefly characterized by an increase in BOLD activity in the absence of externally oriented tasks [@Buckner2008], which is mirrored by a decrease in activity during cognitively demanding tasks compared with less-demanding task conditions or periods of rest [@Laird2009; @Mckiernan2003]. The latter effect has fueled the idea of the DN as a "task-negative" network, although certain self-referential tasks can prompt its activation [@Spreng2012]. Since activity in the DN is related to the availability to engage in introspective thought, its distincitve topographical pattern is often captured by inter-region correlations in resting-state fMRI data [@Buckner2008; @Fox2005; @Greicius2003; @Yeo2011].

  In addition to their role in undirected mentation, mPFC and PCC participate in subjective valuation during economic decision making. Both regions (together with ventral striatum) consistently show greater BOLD activity in response to more highly valued choice prospects and outcomes, relative to prospects and outcomes that are less highly valued [@Bartra2013; @Clithero2014; @Hiser2018; @Kable2007; @Levy2011]. While many cognitive functions can be at least partially separated in mPFC [@DelaVega2016; @Hiser2018; @Kragel2018], coordinate-based meta-analyses of the fMRI literature show that valuation indistinguishably overlaps with the DN in ventral mPFC [@Acikalin2017]. However, subjective valuation and introspective cognition diverge in other brain areas; a region of the striatum is preferentially associated with valuation, whereas bilateral temporoparietal junction is linked with the DN. Within PCC and adjacent precuneus, a more posterior subregion appears DN-specific, whereas a more anterior subregion is implicated in both effects. One possible interpretation of these findings is that mPFC and anterior PCC operate as common nodes within partially overlapping distributed brain systems that subserve valuation and DN-related cognition. This colocalization has prompted theories that these processes share common elements at the psychological level. For example, the two sets of processes might share elements of self-referential cognition, episodic memory, mental simulation, or monitoring and regulation of internal states [@Acikalin2017; @Clithero2014; @Northoff2012; @Reddan2018].

  However, strong conclusions about functional colocalization require consideration of individual-level heterogeneity in topographic patterns of brain activity. A recognized limitation of group averaging and meta-analysis is that the functional topography of individual brains can be misaligned and blurred [@Fedorenko2012; @Guntupalli2018a; @Michalka2015; @Tobyne2018; @Wang2015; @Woo2014], exaggerating the apparent overlap across domains. This concern is especially pronounced in mPFC, which is subject to considerable idiosyncratic cortical folding [@Lopez2019; @Mackey2014; @Zilles2013] and inter-subject functional variability [@Mueller2013]. An alternative approach is to focus on analyses at the individual-participant level. Individual-level analyses of fMRI data have identified idiosyncratic, reliable, and valid patterns of functional organization that would be blurred in aggregative estimates [@Gordon2017; @Gratton2018b; @Laumann2015; @Michalka2015; @Tobyne2018], and subject-specific network arrangements have been found to predict behavioral characteristics [@Kong2018]. Recent work has uncovered fine-grained subdivisions within the DN using both data-driven clustering and individually customized seed-based connectivity analysis [@Braga2017; @Braga2019]. It is therefore possible that the previously reported overlap of the DN with valuation effects can be attributed to low effective spatial resolution, and that the organization of mPFC and PCC would be better understood at the individual level. An important first step, and the goal of the present paper, is to quantify the degree of topographic heterogeneity of the DN within mPFC and PCC in a large sample of individuals. This evaluation of variability during rest will provide a frame of reference for individualized topographical estimation of behavioral effects (such as valuation) in the future, and could provide insights on the extent to which certain task-positive effects attributed to DN regions truly belong to this network [@Spreng2012].

@@

  The outline of our paper is as follows. First, we established a search space by selecting parcels from an established brain atlas [@Glasser2016] that corresponded to previously defined DN and limbic networks on the medial cortical wall [@Yeo2011]. A cortical surface-based meta-analysis of the DN and valuation literatures identified a parcel in PCC that was DN-specific at the aggregate level. We then derived a functional connectivity network of all the surface vertices within the search space for each of `r length(dir(path = './Summary'))` individual resting-state fMRI data sets from the Human Connectome Project [HCP; @VanEssen2012], and used the SP algorithm to subdivide each individual's network into DN and non-DN communities (labeled according to which community included the meta-analytically identified DN-specific parcel in PCC). Focusing on individual vertices in the search space rather than the parcels (as is typical in brain network analyses) allowed us to finely delineate the topographic extent of each community. The resulting communities varied topographically across individuals, while also appearing to follow common organizational principles. Sliding window and test-retest analyses showed that these partitionings were highly similar across scanning days within (but not between) individuals, and that individual-level idiosyncrasy was greater in mPFC. Partitionings obtained from the SP algorithm had higher test-retest reliability than did analogous results from seed-based functional connectivity. We found that the DN community was likely to be located within principal sulci in ventral mPFC and left PCC, but in gyri within superior mPFC and right PCC. Lastly, we describe how the structure of the resulting automatically defined DN and non-DN communities both aligns with and differs from a recently proposed scheme for identifying subdivisions within the DN [@Braga2017; @Braga2019]. Our work highlights the usefulness of estimating brain effects at the individual level in mPFC and PCC, provides a new framework and tool set for future investigations of overlap across cognitive domains, and motivates future re-examination of seemingly inseparable cognitive processes (such as valuation and introspective thought) in these areas.

>>>>

---
title: Spectral partitioning identifies individual heterogeneity in the functional network topography of ventral and anterior medial prefrontal
  cortex
author: "Claudio Toro-Serey^1^, Sean M. Tobyne^1,2^, & Joseph T. McGuire^1^"
csl: apa.csl
output:
  word_document: #default
    reference_docx: Style_reference_draft.docx
  pdf_document:
    highlight: haddock
    keep_tex: no
header-includes:
- \usepackage{setspace}\doublespacing
- \usepackage[left]{lineno}
- \linenumbers
- \usepackage{float}
- \newcommand{\beginsupplement}{ \setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}}
  \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}} }
#subtitle: 'Abbreviated Title: Heterogeneity in the functional topography of mPFC'
bibliography: mPFC_paper.bib
---

@@

1.  Department of Psychological and Brain Sciences and Center for Systems Neuroscience, Boston University, Boston, USA

@@

  Regions of human medial prefrontal cortex (mPFC) and posterior cingulate cortex (PCC) are part of the default network (DN), and additionally are implicated in diverse cognitive functions ranging from autobiographical memory to subjective valuation. Our ability to interpret the apparent co-localization of these effects is constrained by a limited understanding of the individual-level heterogeneity in mPFC/PCC functional organization. Here we used cortical surface-based meta-analysis to identify a parcel in human PCC that was more strongly associated with the DN than with valuation effects. We then used resting-state fMRI data and a data-driven network analysis algorithm, spectral partitioning, to partition mPFC and PCC into "DN" and "non-DN" subdivisions in individual participants (n = 100 from the Human Connectome Project). The spectral partitioning algorithm identified individual-level cortical subdivisions that were reliable across test/retest datasets and varied markedly across individuals, especially in mPFC. Our results point toward a new generation of strategies for assessing whether distinct cognitive functions engage common or distinct mPFC subregions, especially at the individual level.

@@

  Human medial prefrontal cortex (mPFC) and posterior cingulate cortex (PCC) are jointly associated with a diverse set of cognitive processes [@Hiser2018; @Kragel2018], and contain subregions that are part of the brain's default network [DN; @Buckner2008]. DN regions are characterized by a decrease in BOLD activity during externally oriented tasks  that require attention or cognitive control, in comparison with less-demanding task conditions or periods of rest [@Buckner2008; @Laird2009; @Mckiernan2003]. The DN can also be identified on the basis of a distinctive pattern of inter-region correlations in resting-state fMRI data [@Buckner2008; @Fox2005; @Greicius2003; @Yeo2011].

*Many different cognitive task manipulations evoke patterns of brain activity that overlap with DN regions in ventral/anterior mPFC and in PCC. Examples include manipulations of self-referential thinking *(Gusnard et al., 2001;* [@Mitchell2005], memory *(Schacter et al., 2007;*[@Euston2012], affective regulation *(Schiller et al., 2008;*[@Reddan2018], and subjective valuation [@Bartra2013; @Clithero2014; @Kable2007; @Levy2011). Some of these task-related effects are thought to reflect processes integral to the functional role of the DN, such as internally oriented cognition, scene construction, and self-projection [@Hassabis2007]. For other task-related effects, such as subjective valuation (i.e., greater BOLD activity in response to more highly valued choice prospects and outcomes, relative to prospects and outcomes that are less highly valued), the degree of overlap with DN regions is only partial and the reason for the overlap is less obvious [@Acikalin2017]. Insofar as the DN shares a subset of nodes in common with the distributed brain system that supports valuation and other functions, this has potential to inform our theoretical understanding of the cognitive operations involved in those functions [@Northoff2012]. As a result, there is widespread interest in understanding the degree to which DN regions overlap topographically with task-related effects [@Dinicola2019; @Spreng2012].*

  However, strong conclusions about functional colocalization require consideration of individual-level heterogeneity in topographic patterns of brain activity. A recognized limitation of group averaging and meta-analysis is that the functional topography of individual brains can be misaligned and blurred [@Fedorenko2012; @Guntupalli2018a; @Michalka2015; @Tobyne2018; @Wang2015; @Woo2014], exaggerating the apparent overlap across domains. This concern is especially pronounced in ventral mPFC, which is subject to considerable idiosyncratic cortical folding [@Lopez2019; @Mackey2014; @Zilles2013] and inter-subject functional variability [@Mueller2013]. An alternative approach is to focus on analyses at the individual-participant level. Individual-level analyses of fMRI data have identified idiosyncratic, reliable, and valid patterns of functional organization that would be blurred in aggregative estimates [@Gordon2017; @Gratton2018b; @Laumann2015; @Michalka2015; @Tobyne2018], and subject-specific network arrangements have been found to predict behavioral characteristics [@Kong2018]. Recent work has uncovered fine-grained subdivisions within the DN using both data-driven clustering and individually customized seed-based connectivity analysis [@Braga2017; @Braga2019]. It is therefore possible that the previously reported overlap of the DN with task-related effects could, in some cases, be attributed to low effective spatial resolution, and that the organization of mPFC and PCC would be better understood at the individual level. An important first step, and the goal of the present paper, is to quantify the degree of variability in the topography of the DN within mPFC and PCC across a large sample of individuals. The methods we propose for individual-level identification of DN regions have potential to benefit future investigations of the overlap between DN regions and task-related activity.

@@

  The outline of our paper is as follows. First, we defined a search space by selecting parcels from an established brain atlas [@Glasser2016] that corresponded to previously defined DN and limbic networks on the medial cortical wall [@Yeo2011]. A cortical surface-based meta-analysis of the DN and valuation literatures identified a parcel in PCC that was DN-specific at the aggregate level. Valuation was selected as a domain in which task-related activity patterns overlap extensively with DN regions on the medial surface, despite being segregable elsewhere [@Acikalin2017]. We then derived a functional connectivity network of all the surface vertices within the search space for each of `r length(dir(path = './Summary'))` individual resting-state fMRI data sets from the Human Connectome Project [HCP; @VanEssen2012], and used the SP algorithm to subdivide each individual's network into DN and non-DN communities (labeled according to which community included the meta-analytically identified DN-specific parcel in PCC). Focusing on individual vertices in the search space rather than the parcels (as is typical in brain network analyses) allowed us to finely delineate the topographic extent of each community. The resulting communities varied topographically across individuals, while also appearing to follow common organizational principles. Test-retest analyses showed that these partitionings were similar across scanning days within (but not between) individuals, and that individual-level idiosyncrasy was greater in mPFC. Partitionings obtained from the SP algorithm had higher test-retest reliability than did analogous results from seed-based functional connectivity. We found that the DN community was likely to be located within principal sulci in ventral mPFC and left PCC, but in gyri within superior mPFC and right PCC. Lastly, we describe how the structure of the resulting automatically defined DN and non-DN communities both aligns with and differs from a recently proposed scheme for identifying subdivisions within the DN [@Braga2017; @Braga2019]. Our work highlights the usefulness of estimating brain effects at the individual level in mPFC and PCC, provides a new framework and tool set for future investigations of overlap across cognitive domains.
